---
title: "Myeloma2_SM"
author: "SNF"
date: "10/07/20"
output:
  prettydoc::html_pretty:
    theme: Cayman
    highlight: github
editor_options: 
  chunk_output_type: console
---

```{r, warning=FALSE, message=FALSE, warning=FALSE, echo=F}
#devtools::install_github('scfurl/m3addon')
rm(list=ls())
knitr::opts_chunk$set(fig.width=8, fig.height=6,
                      echo=FALSE, warning=FALSE, message=FALSE)
knitr::opts_chunk$set(dev.args=list(bg="transparent"))
ROOT_DIR<-"/fh/fast/furlan_s/user/owalt/exhaustion"
stem<-"mm2"
DATA_DIR <- file.path(ROOT_DIR,  stem, "data")      # SPECIFY HERE
RES_DIR  <- file.path(ROOT_DIR,  stem, "res")     # SPECIFY HERE
RMD_DIR  <- file.path(ROOT_DIR,  stem, "rmd")     # SPECIFY HERE
CDS_DIR <- file.path(ROOT_DIR,  stem, "cds")
FIG_DIR <- file.path(ROOT_DIR,  stem, "figs")
EXT_DIR <- file.path(ROOT_DIR,  stem, "ext_data")

dyn.load('/app/software/ArrayFire/3.8.1/lib64/libaf.so.3')
library(RcppArrayFire)

suppressPackageStartupMessages({
  library(monocle3)
  library(m3addon)
  library(reticulate)
  library(openxlsx)  
  library(dplyr)
  library(Matrix)
  library(ggplot2)
  library(hdf5r)
  library(xfun)
  library(pals)
  library(RColorBrewer)
  library(ggrepel)
  library(stringr)
  library(Seurat)
  library(ArchR)
  library(ggsignif)
  library(viridis)
  library(ggpubr)
  library(data.table)
  library(DESeq2)
  library(pheatmap)
  library(ggrepel)
  library(ComplexHeatmap)
  library(scCustomize)
  library(viewmaster)
  library(tidyverse)
  library(msigdbr)
  library(fgsea)
  library(scRepertoire)
})

set.seed(1234)

#Load colors
h_cols <-rev(brewer.pal(name = "RdYlBu", n = 7))

# Monocle Object
cdsi <-readRDS(file.path(CDS_DIR, "11232021_cdsi.RDS"))

#Seurat Object
mm2<- readRDS(file.path(CDS_DIR, "120621_mm2_seurat.rds"))

fish.integrated<-readRDS(file.path(EXT_DIR, "fish_integrated.rds"))

fish_cols<-c("Gzma+_Tex" ="#16482A","Tgphex" = "#45AC49","Tcm" = "#50191E", "Teff"= "#bc2cdb","Tem" = "#f0821d" ,"Tex_1"= "#3244a8","Tex_cycling"="#2394C4","Tn"="#ffc400","Tpex"="#DC3F32" )

mm2$cluster1<-factor(mm2$cluster1, levels=c("Tn", "Tcm", "Tem", "Tpex", "Teff", "Tex_1", "Tgphex", "Gzma+_Tex", "Tex_cycling"))

sampleID_cols<-c(brewer.oranges(n=3), brewer.blues(n =3), "#d8d9a7")

lcmv_cols<-sfc(20)
names(lcmv_cols)<-levels(factor(fish.integrated$ClusterNames))

rna_cols <- paletteContinuous(n=8)[c(1:3, 6:8)]

set.seed(1234)
```

################# Fig1 ####################
Fig 1A
```{r cells by sortgate}
plot_cells(cdsi[,colData(cdsi)$pop != "FALSE"], color_cells_by = "pop", label_cell_groups = F, cell_size = 0.5)+scale_color_manual(values = c(  "#e3a405","#101b63","#b869bf", "gray"))
```

```{r RNA velocity and cluster}
cdsi@reduce_dim_aux@listData[["UMAP"]]@listData[["model_file"]] <- file.path(CDS_DIR, "201204_Umapmodel")

exprs <- t(assay(cdsi))
col_data <- as.data.frame(colData(cdsi))
row_data <- as.data.frame(rowData(cdsi))
embedding <- reducedDim(cdsi, "UMAP")
```

```{python}
import scanpy as sc

adata_sce = sc.AnnData(X = r.exprs, obs = r.col_data, var = r.row_data)
adata_sce.obsm['umap'] = r.embedding
sc.pl.umap(adata_sce, color='cluster1')

import scvelo as scv
import os
scv.set_figure_params()
filename = os.path.join(r.DATA_DIR, "MM_relapsed/velocyto/MM_relapsed.loom")
ldata = scv.read(filename, cache=True)
adata = scv.utils.merge(adata_sce, ldata)


scv.pp.filter_and_normalize(adata)
scv.pp.moments(adata)
scv.tl.velocity(adata, mode='stochastic')
scv.tl.velocity_graph(adata)
scv.pl.velocity_embedding_stream(adata, basis='umap', color="cluster1", smooth=0.4, min_mass=0, alpha=0.6)
```

```{r Targeted heatmap of genes}
cluster.averages <- AverageExpression(mm2, return.seurat = TRUE, group.by = "cluster1")

tfs<-  c( "Lef1", "Tcf7", "Bach2","Jun", "Tox", "Maf", "Eomes", "Tbx21",  "Nr4a2", "Satb1", "Prdm1", "Irf4", "Batf")
mat <- cluster.averages@assays$RNA@counts[tfs,] %>% as.matrix()
mat<- t(scale(t(mat)))
Heatmap(mat[,c("Tn", "Tcm","Tem","Tpex","Teff", "Tex_1","Gzma+_Tex", "Tgphex",  "Tex_cycling" )], col = h_cols, name = "Expression", row_names_gp = gpar(fontsize = 17), column_names_gp = gpar(fontsize = 17), column_names_rot = 60, cluster_columns = F)

efs<-  c("Gzmb", "Gzma", "Gzmk", "Prf1", "Il10", "Ifng", "Il2", "Fasl", "Ccl2", "Ccl3", "Ccl4", "Ccl5", "Csf1")
mat <- cluster.averages@assays$RNA@counts[efs,] %>% as.matrix()
mat<- t(scale(t(mat)))
Heatmap(mat[,c("Tn", "Tcm","Tem","Tpex","Teff", "Tex_1","Gzma+_Tex", "Tgphex",  "Tex_cycling" )], col = h_cols, name = "Expression", row_names_gp = gpar(fontsize = 17), column_names_gp = gpar(fontsize = 17), column_names_rot = 60, cluster_columns = F)

act_in<-  c( "Mki67", "Entpd1", "Cd38", "Pdcd1", "Cd226", "Lag3", "Ctla4", "Cd244", "Nkg7", "Cd101", "Havcr2")
mat <- cluster.averages@assays$RNA@counts[act_in,] %>% as.matrix()
mat<- t(scale(t(mat)))
Heatmap(mat[,c("Tn", "Tcm","Tem","Tpex","Teff", "Tex-1","Gzma+-Tex", "Maf+-Tex",  "Tex-cycling" )], col = h_cols, name = "Expression", row_names_gp = gpar(fontsize = 17), column_names_gp = gpar(fontsize = 17), column_names_rot = 60, cluster_columns = F)

mem<- c("Cd44",  "Klrg1", "Ccr7", "Slamf6", "Il18r1", "Il7r", "Il2rb", "Sell", "Cxcr5", "Cx3cr1", "Ly6a", "Cd28")
mat <- cluster.averages@assays$RNA@counts[mem,] %>% as.matrix()
mat<- t(scale(t(mat)))
Heatmap(mat[,c("Tn", "Tcm","Tem","Tpex","Teff", "Tex-1","Gzma+-Tex", "Maf+-Tex",  "Tex-cycling" )], col = h_cols, name = "Expression", row_names_gp = gpar(fontsize = 17), column_names_gp = gpar(fontsize = 17), column_names_rot = 60, cluster_columns = F)
```

Supplemental Fig1A
#Feature Plots
```{r}
FeaturePlot_scCustom(mm2, features = "Batf", max.cutoff = "q99", reduction = "RNA_UMAP", pt.size = 0.6, order = T, colors_use =  c("#3361A5", "#1E97F7", "#61C4F4",  "#F0C86D" ,"#EC5724" ,"#A31D1D"), na_color = "gray85")+NoAxes()

FeaturePlot_scCustom(mm2, features = "Ifng", max.cutoff = "q99", reduction = "RNA_UMAP", pt.size = 0.6, order = T, colors_use =  c("#3361A5", "#1E97F7", "#61C4F4",  "#F0C86D" ,"#EC5724" ,"#A31D1D"), na_color = "gray85")+NoAxes()

FeaturePlot_scCustom(mm2, features = "Tcf7", max.cutoff = "q99", reduction = "RNA_UMAP", pt.size = 0.6, order = T, colors_use =  c("#3361A5", "#1E97F7", "#61C4F4",  "#F0C86D" ,"#EC5724" ,"#A31D1D"), na_color = "gray85")+NoAxes()

FeaturePlot_scCustom(mm2, features = "Maf", max.cutoff = "q99", reduction = "RNA_UMAP", pt.size = 0.6, order = T, colors_use =  c("#3361A5", "#1E97F7", "#61C4F4",  "#F0C86D" ,"#EC5724" ,"#A31D1D"), na_color = "gray85")+NoAxes()

FeaturePlot_scCustom(mm2, features = "Gzmb", max.cutoff = "q99", reduction = "RNA_UMAP", pt.size = 0.6, order = T, colors_use =  c("#3361A5", "#1E97F7", "#61C4F4",  "#F0C86D" ,"#EC5724" ,"#A31D1D"), na_color = "gray85")+NoAxes()

FeaturePlot_scCustom(mm2, features = "Gzma", max.cutoff = "q99", reduction = "RNA_UMAP", pt.size = 0.6, order = T, colors_use =  c("#3361A5", "#1E97F7", "#61C4F4",  "#F0C86D" ,"#EC5724" ,"#A31D1D"), na_color = "gray85")+NoAxes()

FeaturePlot_scCustom(mm2, features = "Tox", max.cutoff = "q99", reduction = "RNA_UMAP", pt.size = 0.6, order = T, colors_use =  c("#3361A5", "#1E97F7", "#61C4F4",  "#F0C86D" ,"#EC5724" ,"#A31D1D"), na_color = "gray85")+NoAxes()

FeaturePlot_scCustom(mm2, features = "Prf1", max.cutoff = "q99", reduction = "RNA_UMAP", pt.size = 0.6, order = T, colors_use =  c("#3361A5", "#1E97F7", "#61C4F4",  "#F0C86D" ,"#EC5724" ,"#A31D1D"), na_color = "gray85")+NoAxes()

FeaturePlot_scCustom(mm2, features = "Cx3cr1", max.cutoff = "q99", reduction = "RNA_UMAP", pt.size = 0.6, order = T, colors_use =  c("#3361A5", "#1E97F7", "#61C4F4",  "#F0C86D" ,"#EC5724" ,"#A31D1D"), na_color = "gray85")+NoAxes()

FeaturePlot_scCustom(mm2, features = "Ly6a", max.cutoff = "q99", reduction = "RNA_UMAP", pt.size = 0.6, order = T, colors_use =  c("#3361A5", "#1E97F7", "#61C4F4",  "#F0C86D" ,"#EC5724" ,"#A31D1D"), na_color = "gray85")+NoAxes()

FeaturePlot_scCustom(mm2, features = "Cd28", max.cutoff = "q99", reduction = "RNA_UMAP", pt.size = 0.6, order = T, colors_use =  c("#3361A5", "#1E97F7", "#61C4F4",  "#F0C86D" ,"#EC5724" ,"#A31D1D"), na_color = "gray85")+NoAxes()
```


Fig 1B
#Project Schietinger solid tumor data
```{r}
f<-list.files(file.path(DATA_DIR,"GSE126973_RAW"), full.names = T)

dl<-lapply(f, function(file){
  dat<-fread(file)
  df<-DataFrame(dat[[grep(".bam$", colnames(dat))]], row.names = dat$Geneid, stringsAsFactors = F)
  colnames(df)<-gsub(".merged.bam", "", colnames(dat)[grep(".bam$", colnames(dat))])
  df
  })
mat<-as.matrix(do.call(cbind, dl))
colnames(mat)
g<-GEOquery::getGEO("GSE126973")
l1<-g$GSE126973_series_matrix.txt.gz@phenoData@data$supplementary_file_1 %>% strsplit("_") %>% sapply("[[",2)
l2<-gsub(".gene.counts.txt.gz", "", l1)

pdata<-g$GSE126973_series_matrix.txt.gz@phenoData@data
mat<-mat[,match(colnames(mat), l2)]
pdata$group<-factor(strsplit(l2, "-") %>% sapply("[[", 1))
pdata$sample<-l2
dds<-DESeq2::DESeqDataSetFromMatrix(mat, colData = pdata, design=~group)
sampleConditions <-  factor(make.names(dds$group))
isexpr <- rowSums(edgeR::cpm(counts(dds)) > 1) >= min(table(sampleConditions))
dds <- dds[isexpr,]
dds<-DESeq(dds)
res <- results(dds)
plotMA(res, ylim=c(-10,10))
vsd<-vst(dds)

pd<-project_data(projector = cdsi, projectee = dds, make_pseudo_single_cells = T)

p<-as.data.frame(do.call(rbind, pd[2:1]))
cdata<-colData(dds)[match(p$Type, colnames(dds)),]
cd<-apply(cdata, 1, function(v) v[is.na(v)]<-"scRNAseq")
rownames(cdata)<-rownames(p)
p<-cbind(p, cdata)
p$exp<-p[["biological context:ch1"]]
ggplot(p, aes(x=UMAP1, y=UMAP2, color=group))+geom_point(size=0.4)+SFtheme+scale_color_manual(values=sfc(8), na.value="grey80")+ guides(colour = guide_legend(override.aes = list(size=5)))+facet_wrap(~exp)

ggplot(p, aes(x=UMAP1, y=UMAP2, color=group))+geom_point(size=0.4)+SFtheme+scale_color_manual(values=sfc(8), na.value="grey80")+ guides(colour = guide_legend(override.aes = list(size=5)))

df<-data.frame(DN=cdsi$Log_H1, CD38posCD101neg=cdsi$Log_H2, DP=cdsi$Log_H3)
cdsi$SortedCell<-colnames(df)[apply(df,1,which.max)]

plot_cells(cdsi, color_cells_by = "SortedCell",  label_cell_groups = F, cell_size = 0.7)+SFtheme+scale_color_manual(values=sfc(3))+theme_void()

plot_cells(cdsi, color_cells_by = "cluster1",  label_cell_groups = F, cell_size = 0.7)+SFtheme+scale_color_manual(values=sfc(10))+theme_void()

p$group2<-as.character(p$group)

p$group2[is.na(p$group2)]<-"zcRNAseq"
table(p$group2)
table(subset(p, p$group2 %in% c("N", "zcRNAseq"))$group2)
ggplot(subset(p, p$group2 %in% c("N", "zcRNAseq")), aes(x=UMAP1, y=UMAP2, color=group))+geom_point(size=0.4)+SFtheme+scale_color_manual(values=sfc(8)[3], na.value="grey80")+ guides(colour = guide_legend(override.aes = list(size=5)))+theme_void()

ggplot(subset(p, p$group2 %in% c("TOX","mGFP", "zcRNAseq")), aes(x=UMAP1, y=UMAP2, color=group))+geom_point(size=0.4)+SFtheme+scale_color_manual(values=sfc(8)[c(2,6)], na.value="grey80")+ guides(colour = guide_legend(override.aes = list(size=5)))+theme_void()

ggplot(subset(p, p$group2 %in% c("SV40","OT1", "zcRNAseq")), aes(x=UMAP1, y=UMAP2, color=group))+geom_point(size=0.4)+SFtheme+scale_color_manual(values=sfc(8)[c(4,5)], na.value="grey80")+ guides(colour = guide_legend(override.aes = list(size=5)))+theme_void()

ggplot(subset(p, p$group2 %in% c("KO","WT", "zcRNAseq")), aes(x=UMAP1, y=UMAP2, color=group))+geom_point(size=0.4)+SFtheme+scale_color_manual(values=sfc(8)[c(1,7)], na.value="grey80")+ guides(colour = guide_legend(override.aes = list(size=5)))+theme_void()
```

Fig 1C
#polyclonal data (ring)
```{r}
files<-list.files("../extdata/polyclonal/GSE146609_RAW")
dirs<-paste0("../extdata/polyclonal/GSE146609_RAW/", files)
mat<- Matrix::readMM(dirs[1])

bcs<-read.table("../extdata/polyclonal/GSE146609_barcodes.tsv.gz")
genes<-read.table("../extdata/polyclonal/GSE146609_genes.tsv.gz")

rownames(mat)<-genes$V2 %>% make.unique()
colnames(mat)<-bcs$V1

mat<-mat[which(rowSums(mat)>10),which(colSums(mat)>10)]

dim(mat)

poly<-CreateSeuratObject(counts = mat %>% as.sparse())

# store mitochondrial percentage in object meta data
poly <- PercentageFeatureSet(poly, pattern = "^mt-", col.name = "percent.mt")

poly <- subset(poly, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)

# run sctransform
poly <- SCTransform(poly, method = "glmGamPoi", vars.to.regress = "percent.mt", verbose = T)

# These are now standard steps in the Seurat workflow for visualization and clustering
poly <- RunPCA(poly, verbose = T)
poly <- RunUMAP(poly, dims = 1:15, verbose = T)

poly <- FindNeighbors(poly, dims = 1:15, verbose = T)
poly <- FindClusters(poly, verbose = T)
DimPlot(poly, label = TRUE) + NoLegend()
```
Clusters containing the following cell types were identified using cell type markers: Tumour cells (eGFP), Myeloid cells (Cd68), NK cells (Ncr1), T cells (Cd3e), Neutrophils (Lcn2), and subsets of these groups were identified by markers noted in heatmaps
```{r}
rownames(poly)[grep("eGFP", rownames(poly))]
#tumor
FeaturePlot_scCustom(poly,"gene-eGFP")
#myeloid
FeaturePlot_scCustom(poly,"Cd68")
#NKS
FeaturePlot_scCustom(poly,"Ncr1")
#Neutrophils
FeaturePlot_scCustom(poly,"Lcn2")

#Tcells
FeaturePlot_scCustom(poly,"Cd3e")

poly_t<-poly[,poly$seurat_clusters %in% c(3,7,12,8,1)]

poly_t <- SCTransform(poly_t, method = "glmGamPoi", vars.to.regress = "percent.mt", verbose = T)

# These are now standard steps in the Seurat workflow for visualization and clustering
poly_t <- RunPCA(poly_t, verbose = T)
poly_t <- RunUMAP(poly_t, dims = 1:8, verbose = T)

poly_t <- FindNeighbors(poly_t, dims = 1:8, verbose = T)
poly_t <- FindClusters(poly_t, verbose = T)
DimPlot(poly_t, label = TRUE) + NoLegend()
```

id our signature
```{r}
DefaultAssay(poly_t)<-"SCT"
FeaturePlot_scCustom(poly_t, features = c("Havcr2", "Ifng","Tox", "Batf", "Prf1", "Gzma", "Gzmb",  "Cd4", "Cd8a"), colors_use = rna_cols)

poly_t$seurat_clusters
mm2$cluster1

#get rid of Cd4s
poly_t<-poly_t[,!poly_t$seurat_clusters %in% c(2,5)]

poly_t <- SCTransform(poly_t, method = "glmGamPoi", vars.to.regress = "percent.mt", verbose = T)

# These are now standard steps in the Seurat workflow for visualization and clustering
poly_t <- RunPCA(poly_t, verbose = T)
poly_t <- RunUMAP(poly_t, dims = 1:8, verbose = T)

poly_t <- FindNeighbors(poly_t, dims = 1:8, verbose = T)
poly_t <- FindClusters(poly_t, verbose = T)
DimPlot(poly_t, label = TRUE) + NoLegend()
```

logistic regression
```{r}
library(viewmaster)
library(doMC)

pcds<-seurat_to_monocle3(poly_t)

vg<-common_variant_genes(cdsi, pcds, top_n = 5000)

pcds<-viewmastR::viewmastR(ref_cds = cdsi, query_cds = pcds, ref_celldata_col = "cluster1",selected_genes = vg, verbose=T, FUNC = "softmax_regression", tf_idf = F, learning_rate = 1)

plot_cells(pcds, color_cells_by = "smr_celltype", cell_size = 1)

poly_t$smr_celltype<-pcds$smr_celltype

tab<-table(poly_t$seurat_clusters, poly_t$smr_celltype) %>% as.matrix()
tab<-sweep(tab, 2, colSums(tab), '/')

pdf("../extdata/polyclonal/cd8_pbs_umap.pdf", width = 4, height = 4)
DimPlot(poly_t, label = F, cols = paletteDiscrete(values = levels(factor(poly_t$seurat_clusters)), set = "paired"), pt.size = 1)&NoAxes()
dev.off()

pdf("../extdata/polyclonal/fish_polycloncal_correlation_heatmap.pdf", width = 6, height =6)
pheatmap::pheatmap(tab, cluster_rows = T, cluster_cols=T, color = viridis(10), display_numbers = F)
dev.off()
```


Fig 1D  & Supplemental Fig 1C
#run cca to co-embed fishand lcmv
```{r}
lcmv<-readRDS(file.path(EXT_DIR, "LCMV_Arm_Cl13.Rds"))
DimPlot(lcmv, group.by = "ClusterNames")

# find anchors
anchors <- FindIntegrationAnchors(object.list = list(mm2, lcmv))

# integrate data
fish.integrated <- IntegrateData(anchorset = anchors)

# Run the standard workflow for visualization and clustering
fish.integrated  <- ScaleData(fish.integrated , verbose = FALSE)
fish.integrated  <- RunPCA(fish.integrated , npcs = 50, verbose = FALSE)
ElbowPlot(fish.integrated)
fish.integrated  <- RunUMAP(fish.integrated , reduction = "pca", dims = 1:20, verbose = FALSE)


fish.integrated$cluster1<-mm2$cluster1
fish.integrated$ClusterNames<-lcmv$ClusterNames

is.na(fish.integrated$cluster1) %>% table()
fish.integrated$cluster1[is.na(fish.integrated$cluster1)]<-"X"
fish.integrated$ClusterNames[is.na(fish.integrated$ClusterNames)]<-"X"
table(fish.integrated$cluster1)
table(fish.integrated$ClusterNames)

fish.integrated$cluster1<-factor(fish.integrated$cluster1, levels = c("X", "Gzma+_Tex",  "Tgphex" , "Tcm" ,  "Teff" , "Tem","Tex_1", "Tex_cycling", "Tn", "Tpex"))

table(fish.integrated$ClusterNames)
table(fish.integrated$ClusterNames)
```

```{r}
p1 <- DimPlot(fish.integrated, order =T ,reduction = "umap", group.by = "cluster1", cols =  c("gray80", "#16482A", "#45AC49","#50191E", "#bc2cdb","#f0821d" ,"#3244a8","#2394C4","#ffc400","#DC3F32"), na.value =  "gray85", label = F)&NoAxes() 

p3 <- DimPlot(fish.integrated, reduction = "umap", group.by = "ClusterNames", label = T, label.box = T,  na.value =  "gray85", cols =c(sfc(19), "gray80" ))+scale_fill_manual(values = rep("white", 20))&NoAxes()

p1 +p3
```

```{r id acute and chronic cells}
DimPlot(fish.integrated, reduction = "umap", group.by = "sampleID", label = F, label.box = F,  na.value =  "gray85", cols =sampleID_cols)+scale_fill_manual(values = rep("white", 20))&NoAxes()
```

Fig 1E
#load control data from Tigit paper, re-embed-plot clonality
```{r}
ROOT_DIR<-"/fh/fast/furlan_s/grp/experiments/648_Tigitdata"
DATA_DIR <- file.path(ROOT_DIR, "data")      # SPECIFY HERE
RES_DIR  <- file.path(ROOT_DIR,  "res")     # SPECIFY HERE
RMD_DIR  <- file.path(ROOT_DIR,"rmd")     # SPECIFY HERE
CDS_DIR <- file.path(ROOT_DIR,   "cds")

cds<-readRDS(file.path(CDS_DIR, "tigit_cds.rds"))
tig<-monocle3_to_seurat(cds)
tig<-FindVariableFeatures(tig)
tig<-ScaleData(tig)

samps<-list.files(DATA_DIR)[2:5]
folders<-list.files(file.path(DATA_DIR), full.names = T)
colnames(cds)
tcrl<-lapply(samps, function(x){
  tcr<-read.csv(file.path(DATA_DIR, x, "filtered_contig_annotations.csv"))
  tcr
})
#edit vdj files so barcodes match the cell colnames
tcrl2 <- lapply(1:length(tcrl), function(x){
  met<-tcrl[[x]]
  met$barcode<-gsub("-1", paste0("-",x), met$barcode)
  met
})

combined <- combineTCR(tcrl, samples = samps, cells ="T-AB")
combined$Gp1_VDJ$barcode<-gsub(".*_", "", combined$Gp1_VDJ$barcode)
combined$Gp2_VDJ$barcode<-gsub(".*_", "", combined$Gp2_VDJ$barcode)
combined$Gp3_VDJ$barcode<-gsub(".*_", "", combined$Gp3_VDJ$barcode)
combined$Gp4_VDJ$barcode<-gsub(".*_", "", combined$Gp4_VDJ$barcode)

combined$Gp1_VDJ$barcode<-gsub("Gp1_VDJ_", "", combined$Gp1_VDJ$barcode)
combined$Gp2_VDJ$barcode<-gsub("Gp2_VDJ_", "", combined$Gp2_VDJ$barcode)
combined$Gp2_VDJ$barcode<-gsub("1", "2", combined$Gp2_VDJ$barcode)
combined$Gp3_VDJ$barcode<-gsub("Gp3_VDJ_", "", combined$Gp3_VDJ$barcode)
combined$Gp3_VDJ$barcode<-gsub("1", "3", combined$Gp3_VDJ$barcode)
combined$Gp4_VDJ$barcode<-gsub("Gp4_VDJ_", "", combined$Gp4_VDJ$barcode)
combined$Gp4_VDJ$barcode<-gsub("1", "4", combined$Gp4_VDJ$barcode)

seu <- combineExpression(combined, tig, 
                  cloneCall="aa", 
                  proportion = FALSE, 
                  cloneTypes=c(Single=1, Small=5, Medium=20, Large=100, Hyperexpanded=500))

con<-seu[, seu$sample_id == "648_Tigit_Gp4"]
con <-NormalizeData(con)
con<-FindVariableFeatures(con)
con<-ScaleData(con)
con <- SCTransform(con, method = "glmGamPoi", vars.to.regress = "percent.mito")
con <- RunPCA(con, features = VariableFeatures(con))
ElbowPlot(con, ndims = 30)
con<-RunUMAP(con, dims = 1:20, verbose = T, reduction.name = "SCT_UMAP")
DimPlot(con, reduction = "SCT_UMAP", group.by = "sample_id")


slot(con, "meta.data")$cloneType <- factor(slot(con, "meta.data")$cloneType, 
                levels = c("Hyperexpanded (100 < X <= 500)", 
                           "Large (20 < X <= 100)", 
                            "Medium (5 < X <= 20)", 
                            "Small (1 < X <= 5)", 
                            "Single (0 < X <= 1)", NA))

con$cloneType<-as.character(con$cloneType)

con$cloneType[which(is.na(con$cloneType))]<-"NA"
con$cloneType

con$cloneType<-factor(con$cloneType, levels= rev(c("Hyperexpanded (100 < X <= 500)", "Large (20 < X <= 100)", "Medium (5 < X <= 20)", "Small (1 < X <= 5)",  "Single (0 < X <= 1)", "NA")))

DimPlot(con, group.by = "cloneType", order = T)+scale_color_manual(values =c("gray80",rev(brewer.spectral(n=5))))&NoAxes()
```

#use cdsi to annotate control tigit data
```{r}
plot_cells(cdsi, color_cells_by = "cluster1")

con<-seurat_to_monocle3(con, seu_rd = "SCT_UMAP")
cols <- c( "Gzma+_Tex" = "#16482A","Tgphex"="#45AC49","Tcm" ="#50191E","Teff"= "#bc2cdb","Tem"="#f0821d" , "Tex_1"=  "#3244a8", "Tex_cycling"= "#2394C4", "Tn"= "#ffc400", "Tpex"= "#DC3F32")

vg<-common_variant_genes(cdsi, con, top_n = 1000)
cds<-viewmaster(ref_cds  = cdsi, query_cds =con, ref_celldata_col = "cluster1", selected_genes = vg, verbose=T, FUNC = "softmax_regression", tf_idf = F, iterations = 1500)

cds$smr_celltype<-factor(cds$smr_celltype, levels = levels(factor(cds$smr_celltype))[c(1:2, 4:10,3)])

con$smr_celltype<-cds$smr_celltype
DimPlot(con, group.by = "smr_celltype", order = F)+scale_color_manual(values = cols)&NoAxes()
```

Fig 1F
#Braun et al Progessive immune dysfunction
```{r}
genes<-read.csv(file = "../gene_sets/Wu_progressive_immune.csv")
genes$Cluster_Name %>% table()

tex<-genes[grep("Terminally", genes$Cluster_Name),] %>% group_by(Cluster_Name) %>% filter(logFC > 0.25)
t_genes<-str_to_title(tex$gene)

mm2<-AddModuleScore(mm2, features = list(t_genes), name = "Braun")

FeaturePlot_scCustom(mm2, features = "Braun1", pt.size = 0.7)+scale_color_gradientn(colors = paletteContinuous(n=8)[c(1:3, 6:8)])&NoAxes()

cdsi$braun1<-mm2$Braun1

plot_cells(cdsi[,order(cdsi$braun1)], color_cells_by = "braun1", cell_size = 1)
```

Fig 1G
#Human tumor TCR gene set
```{r}
h_genes<-read.csv(file = "../gene_sets/Human neoantigen gene set.csv")
h_genes<-str_to_title(h_genes$NeoTCR8.ALL)

mm2<-AddModuleScore(mm2, features = list(h_genes), name = "Human")

VlnPlot(mm2, features = "Human1", cols =cols, group.by = "cluster1" , pt.size = 0)+ggtitle(label="neoantigen human geneset")
```


